---
title: "R Bingo"
author: "Katie August"
date: "20/10/2020"
output: html_document
---


Here is the code for a game of bingo! One player is assigned the bingo caller role, and they use the code in the fist section to randomly sample numbers from 1 to 99. The bingo caller then slowly calls out the numbers in the order they are generated, while the players score these numbers off of their own cards.

## Bingo Caller Code
```{r}
## Create a sqeunce from 1 to 99 in steps of 1
numbers <- seq(1, 99, 1)

## Code to randomly draw numbers. Replace = FALSE so that a number can only be drawn once
sample(numbers, 99, replace=FALSE)

```

## Bingo Player Card Creator
```{r table, results = "asis"}

## Create vectors to sample for each coloumn of your bingo card
a <- seq(1, 19, 1)
b <- seq(20, 39, 1)
c <- seq(40, 59, 1)
d <- seq(60, 79, 1)
e <- seq(80, 99, 1)

## Take 3 samples of each card column vectors and sort them into ascending order
as <- sort(sample(a, 3, replace = FALSE))
bs <- sort(sample(b, 3, replace = FALSE))
cs <- sort(sample(c, 3, replace = FALSE))
ds <- sort(sample(d, 3, replace = FALSE))
es <- sort(sample(e, 3, replace = FALSE))

## Create a dataframe from these sampled vectors to produce the card
card <- data.frame(as, bs, cs, ds, es)

## Make the bingo card look nice and easier to read
### If you don't have these packages installed, use install.packages("package name")
library(knitr)
library(kableExtra)
library(magrittr)

kable(card, col.names = NULL) %>%
kable_styling("striped", full_width = T, font_size = 36) %>%
column_spec(1:5, bold = T) %>%
row_spec(1:3, bold = T, color = "white", background = "#D7261E")

```
To play, just listen out to the numbers being called by the bingo caller and take note. Once you have a line, and if you are the first person to have a line, shout "Line!". Well done, you are on your way to winning. Once all your numbers have been called, shout "Bingo!". Congratulations, you have won the game!

To play again, switch around who is the bingo caller, and then rerun all of the code to get a new card. Enjoy!

Once you've finished playing, feel free to have a play around with the code to see what other card designs you can create.