---
title: "Other games in R"
author: "Katie August"
date: "31/10/2020"
output: html_document
---

Here you can find code for some solo games that can be played within R. Have fun, and let the group know any high scores!

## Required packages
```{r}
# Don't forget you may need to install these packages using install.packages("package name")

library(fun)
```

## Minesweeper

Minesweeper is a single-player puzzle video game. The objective of the game is to clear a rectangular board containing hidden "mines" or bombs without detonating any of them, with help from clues about the number of neighboring mines in each field.

Controls: Click the left mouse button to dig in an area, and right button to mark or unmark the area with flags.
```{r}
# Run this line to open an interactive graphics window
X11(type="Xlib")

# This line will open the game in the graphics window
mine_sweeper(width = 10, height = 10, mines = 20, cheat = FALSE)

```


## Lights out

By default, the white squares in the plot denote the lights that are on, and black ones for the off.
When you click on a light, this light as well as the four neighbors will switch theirs status. Your
mission is to close all the lights

```{r}
# Run this line to open an interactive graphics window
X11(type="Xlib")

# This line will open the game in the graphics window
lights_out(
width = 5,
height = 5,
steps = 3,
cheat = FALSE,
col.off = "black",
col.on = "white",
col.frame = "lightblue",
seed = NULL
)

```

## Sliding puzzle

The aim of this game is to get the numbers 1-8 in order going lowest in the top left, going up row by row.

```{r}
# Run this line to open an interactive graphics window
X11(type="Xlib")

# This line will open the game in the graphics window

sliding_puzzle(size = c(3, 3), bg = "lightblue", z = NULL)
```

